---
title: "Exercise 8"
date: 09/20/2017
author: "Andy Tse"
output: html_document
runtime: shiny
---

## Loading the Packages
```{r, echo = FALSE}
library(markdown)
library(knitr)
library(shiny)
```


## Using the Three Input Arguments
```{r}
inputPanel(
  
  sliderInput("n", label = "Random Support Values:",
              min = 10, max = 1000, value = 100, step = 10),
  sliderInput("mu", label = "Mean:",
              min = -2, max = 2, value = 0, step = 0.1),
  sliderInput("sigma", label = "Standard Deviation:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

xmin <- -5
xmax <- 5
x <- seq(xmin,xmax,length.out=100)

renderPlot({
  y <- dnorm(x, input$mu, input$sigma)
  plot(x,y, type="l", 
       main = paste0("Normal Distribution X~N(",input$mu,",",input$sigma,")"),
       ylab="Density",
       xlim = c(-5, 5), 
       ylim = c(0,1))

})
```

## Histogram
```{r}
renderPlot({
  rnx <- rnorm(input$n, input$mu, input$sigma)
  hist(rnx, probability = TRUE, xlab = "X Prob", main = "Histogram of RN", xlim = c(-4,4), ylim = c(0,1))
})

```


## Description

The input for the "n" values on the slider gives the random samples of "n" on how it is analyzed. The more the sample size is given, the data would be less skewed for the histogram distribution. However, when the mu and sigma are both set at zero, the normal distribution plot is symmetric. However, for the mu and sigma being set to the max, it is heading more to the center on both directions. When the mean and mu goes all the way to the left, it is headed to negative distributions. 
